### 2013-12-16 research/paper/Theory/mathematical_model.tex
Committing what I think is my last bit before submission. Kasun still needs to fix some stuff up, there are notes in the document to him

The base quasi-steady state (QSS) model was first developed by \citet{Parkinson1964} for a square cross section. The equation of motion of the body  is given by 
\begin{equation}
\label{equationofmotion}
(m+m_a)\ddot{y}+c\dot{y}+ky=F_y,
\end{equation}
\JL{KASUN: why do you have an added mass term here? This implies that you have decomposed the lift force into components. I think you should remove it here and from the rest of the document.}
where the forcing term $F_y$ is given by
\begin{equation}
\label{lift equation}
F_y=\frac{1}{2}\rho U^2\mathcal{A}C_y.
\end{equation}

@@

In the QSS model, it is assumed that the force on the body at a given instantaneous angle of attack $\theta$ (defined in figure \ref{fig:setup_1}) is the same as the mean force on a static body at the same angle of attack. The instantaneous value of $C_y$ is therefore determined by an interpolating polynomial based on the lift and drag data  for flow over a stationary body at various $\theta$. Using the relationship between $\theta$ and the instantaneous transverse velocity of the body $\dot{y}$ shown in figure \ref{fig:setup_1}, $C_y$ can be written as a function of $\dot{y}$. The order of the interpolation polynomial used to define this function has varied from study to study. For  example a $7^{th}$ order polynomial was used in \cite{Parkinson1964} and $3^{rd}$ order polynomial was used in \cite{Barrero-Gil2009}. \cite{Ng2005} concluded that using a $7^{th}$ order polynomial is sufficient and a polynomial higher than that of $7^{th}$ order doesn't provides a significantly better result. Thus a $7 ^{th}$ order interpolating polynomial is used in this present study. As a result, $C_y(\dot{y})$ is defined as
\begin{equation}
\label{cy ploynomial}
C_y(\theta)=a_1\left(\frac{\dot{y}}{U}\right)+a_3\left(\frac{\dot{y}}{U}\right)^3+a_5\left(\frac{\dot{y}}{U}\right)^5+a_7\left(\frac{\dot{y}}{U}\right)^7.
\end{equation}

@@

\begin{equation}
\label{final_equation_motion}
(m{+}m_a)\ddot{y}{+}c\dot{y}{+}ky{=}\frac{1}{2}\rho U^2 \mathcal  {A} \Bigg(a_1\left(\frac{\dot{y}}{U}\right){+}a_3\left(\frac{\dot{y}}{U}\right)^3{+}a_5\left(\frac{\dot{y}}{U}\right)^5{+}a_7\left(\frac{\dot{y}}{U}\right)^7 \Bigg){+} F_0\sin{(\omega_s t)}.
\end{equation}

This equation could be solved using standard time integration methods. In  this study the fourth-order Runge-Kutta scheme built in to the MATLAB routine `ode45' was generally used to obtain the solutions. Some low mass ratio cases used a solver modified for stiff problems, built into the `ode15s' routine in MATLAB.

@@

For the low \reynoldsnumber\ tests, \reynoldsnumber=165 was maintained as it was pointed out by \citet{Sheard2009} and \citet{Tong2008} that the three-dimensional transition for a square cylinder occurs at approximately \reynoldsnumber=160. $F_0$ was kept at $0.4937$ which was obtained by scaling the value used by \citet{Joly2012} with the amplitude ratios of the lift forces obtained at the different Reynolds numbers.

The angular vortex shedding frequency $\omega_s$, was set to $0.98$ which was obtained by performing a power spectral analysis of the stationary data at $0^\circ$. Stationary $C_y$ data were obtained at different angles of attack ranging from $0^\circ$ to $16^\circ$. The average power was obtained by using Eq.\eqref{power}, and the averaging was done over no less than 20 galloping periods. Predictions of power output at \reynoldsnumber=22300 were obtained using the coefficients for curve fitting $C_y$ (Table (\ref{table:cy-coefficients})) from \citet{Parkinson1964}, in order to provide a comparison between high and low Reynolds numbers. The mass ration $m^*$ was kept at 1163 for \reynoldsnumber=22300 (Similar as \citet{Parkinson1964}) and $m^*=20$ for \reynoldsnumber=165. These parameters were used throughout this study unless otherwise specified.

The stationary data and the fluid-structure interaction (FSI) data were obtained using a high-order spectral element routine to simulate the two-dimensional laminar flow.  Simulations involving fluid structure interaction (FSI) were used to provide additional validation of the QSS model. The inlet was placed $20D$ while the outlet situated $60D$ away from the centroid of the body. The side boundaries were placed $20D$ away from the centroid of the body where $D$ was kept as unity throughout this study. The Navier--Stokes equations were solved in an accelerated frame of reference attached to the moving body along with the body equation of motion given in equation \ref{equationofmotion}. A three-step time splitting scheme together with high-order Lagrangian polynomials were used to obtain the solution. The details of the method can be found in \cite{Thompson2006,Thompson1996a}. This code has been very well validated in a variety of fluid-structure interaction problems \citep{Leontini2007a,Griffith2011,Leontini2011,Leontini2013}.

 The computational domain consists of 690 quadrilateral macro elements where majority of the elements were concentrated near the square section. A freestream condition was given to the inlet, top and bottom boundaries and the normal velocity gradient was set to zero at the outlet. A convergence study was performed by changing the order of the polynomial ($p$-refinement) at $U^*=40$ and \reynoldsnumber=$165$. A $9^{th}$ order polynomial together with a time step of $\Delta tU/D=0.001$ was sufficient to ensure an accuracy of $2\%$ with regards to amplitude of oscillation.

>>>>

The base quasi-steady state (QSS) model was first developed by \citet{Parkinson1964} for a square cross section. The equation of motion of the body is given by 
\begin{equation}
\label{equationofmotion}
m\ddot{y}+c\dot{y}+ky=F_y,
\end{equation}
where the forcing term $F_y$ is given by
\begin{equation}
\label{lift equation}
F_y=\frac{1}{2}\rho U^2\mathcal{A}C_y.
\end{equation}
\citet{lighthill1986} showed that for systems oscillating in fluid, it is sometimes useful to decompose the fluid forces into components that are in and out of phase with the body acceleration. The component in phase with the acceleration effectively adds to the inertia or effective mass of the system. Therefore, an added mass term, $m_a$, can be added to the system mass. For consistency with previous studies such as \citet{Joly2012}, a value of $m_a=3.5$ has been used here.

@@

In the QSS model, it is assumed that the force on the body at a given instantaneous incident angle $\theta$ (defined in figure \ref{fig:setup_1}) is the same as the mean force on a static body at the same incident angle, or angle of attack. The instantaneous value of $C_y$ is therefore determined by an interpolating polynomial based on the lift data for flow over a stationary body at various $\theta$. Using the relationship between $\theta$ and the instantaneous transverse velocity of the body $\dot{y}$ shown in figure \ref{fig:setup_1}, $C_y$ can be written as a function of $\dot{y}$. The order of the interpolation polynomial used to define this function has varied from study to study. For  example a $7^{th}$ order polynomial was used in \cite{Parkinson1964} and $3^{rd}$ order polynomial was used in \cite{Barrero-Gil2009}. \cite{Ng2005} concluded that using a $7^{th}$ order polynomial is sufficient and a polynomial higher than that of $7^{th}$ order doesn't provides a significantly better result. Thus a $7 ^{th}$ order interpolating polynomial is used in this present study. As a result, $C_y(\dot{y})$ is defined as
\begin{equation}
\label{cy ploynomial}
C_y(\theta)=a_1\left(\frac{\dot{y}}{U}\right)+a_3\left(\frac{\dot{y}}{U}\right)^3+a_5\left(\frac{\dot{y}}{U}\right)^5+a_7\left(\frac{\dot{y}}{U}\right)^7.
\end{equation}

@@

\begin{equation}
\label{final_equation_motion}
m\ddot{y}{+}c\dot{y}{+}ky{=}\frac{1}{2}\rho U^2 \mathcal  {A} \Bigg(a_1\left(\frac{\dot{y}}{U}\right){+}a_3\left(\frac{\dot{y}}{U}\right)^3{+}a_5\left(\frac{\dot{y}}{U}\right)^5{+}a_7\left(\frac{\dot{y}}{U}\right)^7 \Bigg){+} F_0\sin{(\omega_s t)}.
\end{equation}

This equation can be solved using standard time integration methods. In this study the fourth-order Runge-Kutta scheme built in to the MATLAB routine `ode45' was generally used to obtain the solutions. Some low mass ratio cases used a solver modified for stiff problems, built into the `ode15s' routine in MATLAB.

@@

For the low \reynoldsnumber\ tests, $\reynoldsnumber=165$ was maintained as it was pointed out by \citet{Sheard2009} and \citet{Tong2008} that the three-dimensional transition for a square cylinder occurs at approximately \reynoldsnumber=160. $F_0$ was kept at $0.4937$ which was obtained by scaling the value used by \citet{Joly2012} with the amplitude ratios of the lift forces obtained at the different Reynolds numbers.

The angular vortex shedding frequency $\omega_s$, was set to $0.98$ which was obtained by performing a power spectral analysis of the stationary data at $0^\circ$. Stationary $C_y$ data were obtained at different angles of attack ranging from $0^\circ$ to $16^\circ$. The average power was obtained by using equation \ref{power}, and the averaging was done over no less than 20 galloping periods. Predictions of power output at $\reynoldsnumber=22300$ were obtained using the coefficients for curve fitting $C_y$ (Table (\ref{table:cy-coefficients})) from \citet{Parkinson1964}, in order to provide a comparison between high and low Reynolds numbers. The mass ratio $m^*$ was kept at 1163 for $\reynoldsnumber=22300$ (Similar to \citet{Parkinson1964}) and $m^*=20$ for \reynoldsnumber=165. These parameters were used throughout this study unless otherwise specified.

The stationary data and the fluid-structure interaction (FSI) data were obtained using a high-order spectral element routine to simulate the two-dimensional laminar flow.  Simulations involving fluid structure interaction (FSI) were used to provide additional validation of the QSS model. The inlet was placed $20D$ while the outlet situated $60D$ away from the centroid of the body. The side boundaries were placed $20D$ away from the centroid of the body where $D$ was kept as unity throughout this study. The Navier--Stokes equations were solved in an accelerated frame of reference attached to the moving body along with the body equation of motion given in equation \ref{equationofmotion}. A three-step time splitting scheme together with high-order Lagrangian polynomials were used to obtain the solution. The details of the method can be found in \citet{Thompson2006,Thompson1996a}. This code has been very well validated in a variety of fluid-structure interaction problems \citep{Leontini2007a,Griffith2011,Leontini2011,Leontini2013}.

The computational domain consists of 690 quadrilateral macro elements where the majority of the elements were concentrated near the square section. A freestream condition was given to the inlet, top and bottom boundaries and the normal velocity gradient was set to zero at the outlet. A convergence study was performed by changing the order of the polynomial ($p$-refinement) at $U^*=40$ and $\reynoldsnumber=165$. A $9^{th}$ order polynomial together with a time step of $\Delta tU/D=0.001$ was sufficient to ensure an accuracy of $2\%$ with regards to amplitude of oscillation.

### 2013-12-16 research/paper/figure/m_star_collapsed.tex
Committing what I think is my last bit before submission. Kasun still needs to fix some stuff up, there are notes in the document to him

\setlength{\unitlength}{\textwidth}

  \begin{picture}(1,0.3)(0.0,0.45)

       \put(0.025,0.5){\includegraphics[width=0.5\unitlength]{../FnP/gnuplot/mean_power_collapsed_mstar.eps}}
       \put(0.495,0.5){\includegraphics[width=0.5\unitlength]{../FnP/gnuplot/mean_power_collapsed_noshed_mstar.eps}}

@@

        \put(0.23,0.48){$\displaystyle\frac{c}{\rho\mathcal{A}U}$} 	
        \put(0.73,0.48){$\displaystyle\frac{c}{\rho\mathcal{A}U}$}

@@

  \caption{Mean power as a function of damping factor. Data are presented at $m^*=10$ (\ding{108}), $m^*=20$ (\ding{83}), $m^*=40$ (\ding{115}), $m^*=60$ (+) at Re 165 (a) with and (b) without the shedding term in Eq.\ref{final_equation_motion}. A reduction of maximum mean power can be observed when $m^*<40$ with shedding while the maximum power is essentially independent of $m^*$ when shedding is disregarded.}
    \label{fig:m_star_collapsed}
\end{figure}

>>>>

  \setlength{\unitlength}{\textwidth}
  \begin{picture}(1,0.3)(0.0,0.45)
    \put(0.025,0.5){\includegraphics[width=0.5\unitlength]{../FnP/gnuplot/mean_power_collapsed_mstar.eps}}
    \put(0.495,0.5){\includegraphics[width=0.5\unitlength]{../FnP/gnuplot/mean_power_collapsed_noshed_mstar.eps}}

@@

    % \put(0.23,0.48){$\displaystyle\frac{c}{\rho\mathcal{A}U}$} 	
    % \put(0.73,0.48){$\displaystyle\frac{c}{\rho\mathcal{A}U}$}

    \put(0.28,0.48){\massdamp} 	
    \put(0.78,0.48){\massdamp}

@@

  \caption{Mean power as a function of damping factor. Data are presented at $m^*=10$ (\ding{108}), $m^*=20$ (\ding{83}), $m^*=40$ (\ding{115}), $m^*=60$ (+) at Re 165 (a) with and (b) without the shedding term in equation \ref{final_equation_motion}. A reduction of maximum mean power can be observed when $m^*<40$ with shedding while the maximum power is essentially independent of $m^*$ when shedding is disregarded.}
    \label{fig:m_star_collapsed}
\end{figure}

### 2013-12-16 research/paper/nomenclature.tex
Committing what I think is my last bit before submission. Kasun still needs to fix some stuff up, there are notes in the document to him

\begin{tabular}{ll}
$a_1,a_3,a_5,a_7$ & coefficients of the polynomial to determine $C_y$ \\ 
$C_y$ & instantaneous force coefficient normal to the induced velocity \\
$F_y$ & force due to $C_y$  \\ 
$\rho$ & fluid density  \\
$m$ & mass of the body \\
$m_a$ & added mass \\
$c$ & damping constant/damping factor \\
$k$ & spring constant \\
$U_i$ & induced velocity \\
$\theta$ & induced angle of incidence \\
$U$ & freestream velocity \\
$y,\dot{y},\ddot{y}$ & transverse displacement, velocity and acceleration   \\
$A$ & displacement amplitude\\
$\mathcal{A}$ & cross sectional area\\ 
$F_0$& amplitude of the oscillatory force due to shedding \\
$\omega_s$& vortex shedding frequency \\
$t$ & time \\
$P_{mean}$& mean power \\
$f=\frac{1}{2\pi}\sqrt{\frac{k}{m}}$& natural frequency of the system \\
$\omega_n= 2 \pi f$& natural frequency of the system  \\
$D$ & characteristic length of the body  \\
$m^*=\frac{m}{\rho \times \text{Volume of the body}}$& mass ratio \\
$U^*=\frac{U}{f \times D}$& reduced velocity  \\
$\zeta= \frac{c}{2 m \omega_n}$& damping ratio \\
$P_t$   & power transferred to the body by the fluid \\
$P_d$& power dissipated due to mechanical damping  \\
$Re$& Reynolds number  \\
$\theta= \tan^{-1}{(\frac{\dot{y}}{U})}$& instantaneous angle \\
& \\
& \\
\end{tabular}

>>>>

\begin{tabular}{ll}
$a_1,a_3,a_5,a_7$ & coefficients of the polynomial to determine $C_y$ \\ 
$A$ & displacement amplitude\\
$c$ & damping constant \\
$D$ & characteristic length (side length) of the cross section of the body \\
$f=\sqrt{k/m}/2\pi$ & natural frequency of the system \\
$F_y$ & instantaneous force normal to the flow \\ 
$F_0$& amplitude of the oscillatory force due to vortex shedding \\
$k$ & spring constant \\
$m$ & mass of the body \\
$m_a$ & added mass \\
$P_d$ & power dissipated due to mechanical damping  \\
$P_{mean}$ & mean power \\
$P_t$   & power transferred to the body by the fluid \\
$t$ & time \\
$U$ & freestream velocity \\
$U_i$ & Induced velocity \\
$y,\dot{y},\ddot{y}$ & transverse displacement, velocity and acceleration of the body \\
$\mathcal{A}=DL$ & frontal area of the body\\ 
$\rho$ & fluid density  \\
$\omega_n= 2 \pi f$ & natural angular frequency of the system  \\
$\omega_s$ & vortex shedding angular frequency \\
$\cstar=cD/mU$ & non-dimensionalised damping factor \\
$C_y=F_y/0.5\rho U^2DL$ & normal (lift) force coefficient \\
$m^*=m/\rho D^2L$ & mass ratio \\
$Re$ & Reynolds number  \\
$U^*=U/fD$ & reduced velocity  \\
$Y=y/D$ & non-dimensional transverse displacement \\
$\dot{Y}=m^*\dot{y}/a_1U$ & non-dimensional transverse velocity \\
$\ddot{Y}=m^{*2}D/a_1^2U^2$ & non-dimensional transverse acceleration \\
$\Gamma_1 = 4\pi^2m^{*2}/U^{*2}a_1^2$ & First dimensionless group arising from linearised, non-dimensionalised equation of motion\\
$\Gamma_2 = c^*m^*/a_1$ & Second dimensionless group arising from linearised, non-dimensionalised equation of motion\\
$\zeta= c/2 m \omega_n$ & damping ratio \\
$\theta= \tan^{-1}{(\dot{y}/U)}$ & instantaneous angle of incidence (angle of attack)\\
$\massstiff =  4\pi^2m^{*2}/U^{*2}$ & Combined mass-stiffness parameter\\
$\massdamp = c^*m^*$ & Combined mass-damping parameter\\
\end{tabular}
